A Spatial AR System for Axis-aligned Augmentation of Planar Scenes in Industrial Settings

### 2021-07-09 paper/elsarticle/elsarticle-template.tex
paper

\title{A Spatial AR System for Wide-area Axis-aligned Augmentation of Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

@@

Using the $X$- and $Y$-axes of an image of the scene as a proxy, our contribution is to propose a spatial AR system that produces the effect of keystone correction analytically, and in a manner aligning the axes of the augmentation with those of the proxy image. We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera. This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction. Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications.

@@

Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view. This is because the scene point `illuminated' by a pixel in the projector's image plane is given by intersecting its corresponding ray with the geometry of the scene surface. To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect. Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene. These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.

@@

\begin{figure}
    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
    \caption{Scene plane (gray) obtained via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red). Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the camera (frustum of the camera in black, with up vector in red).}
    \label{fig:3d}
\end{figure}

@@

\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.

@@

If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane. If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired distortion. Projecting an image distorted in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the undistorted image to the scene plane from the viewpoint of the virtual projector. This distortion can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.

\begin{figure}
    \centering
    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/3.png}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3.png}}}
    \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane. The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to render them parallel with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected. This placement is produced according to a small handful of steps. First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}). Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); then, we (iii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector. Finally, we (iv) rotate the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}). The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions. Optionally, we additionally adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.

@@

We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. We contrast our approach with two variants of a baseline approach involving manual keystone correction (`Baseline A' and `Baseline B'), by aiming to place an image for each location aligned with the principal axes of the floorspace, absent projective distortions, and with the same dimensions (30 cm $\times{}$ 20 cm). Using our approach to carry out keystone correction analytically, we carried out all steps within a time budget of 15 min. Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector. The second variant of the baseline approach (`Baseline B') is to again use the bundled software to carry out keystone correction manually, but to record how much time is necessary to produce augmentations within a mean tolerance of XX. All experiments were carried out jointly by the same experienced technician and assistant.

The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}, bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset. In addition, we used a downward-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The setup was mounted on the ceiling of the Pilotfabrik. The floorspace used for our experiments measured dimensions of ca.\ X~m~$\times$~Y~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.

\paragraph{Our approach} We begin by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's pose relative to the projector-steerable mirror system remains fixed.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module. Separately, for each of the 15 target locations, we produce a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera. We then fed these images alongside the output of the camera calibration module to our projector calibration module. For each of the target locations, the steerable mirror is made to point to that location, the circles pattern image is projected to the scene plane, an image is acquired using the camera, and the location is registered in the MDC-X software as a preset. The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}). Next, we warp the images to be projected to the respective locations using their corresponding homography, using a third dedicated module. These warped images are finally imported into the MDC-X software and associated with their respective location presets.

The total amount of time to carry out all the above steps amounted to ca.\ 15 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 4 min going to that of projector calibration images. The remainder of the time was spent running our modules or working with the MDC-X software.

>>>>

\title{A Spatial AR System for Axis-aligned Augmentation of Planar Scenes in Industrial Settings} %\title{Elsevier \LaTeX\ template\tnoteref{mytitlenote}}
%\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

@@

Our contribution is to propose a spatial AR system that produces the effect of keystone correction analytically, and---using the $X$- and $Y$-axes of an image of the scene acquired by a downwards-facing camera as a proxy---in a manner aligning the axes of the augmentation with those of the proxy image. We achieve this by distorting the image to be projected using a plane-induced homography computed to produce the effect of projecting the image not from the actual projector viewpoint, but in accordance with the viewpoint of a \textit{virtual} projector (i) facing directly downwards to the scene plane and (ii) rotated to place the axes of the image plane of the virtual projector in line with those of the camera. This facilitates placement of augmentations by intuitively placing them in accordance with the axes of an image of the scene, and eliminates the need for manual keystone correction. Moreover, a consequence of our approach is that our system is able to handle a projector equipped with a steerable mirror (without need for explicitly modeling the action of the steerable mirror on the projector), thereby enabling wide-area factory floor applications.

@@

Correcting for projective distortions of the sort outlined in Section~\ref{sec:intro} can be achieved by modeling the manner in which the respective rays through the pixels of the projector's image plane fan out into the scene (i.e., by `calibrating' the projector) and the geometry of the scene itself (i.e., by recovering the scene plane relative to the projector) within at least the projector's field of view. This is because the scene point `illuminated' by a pixel in the projector's image plane can be recovered by intersecting its corresponding ray with the geometry of the scene surface. To model this interaction, we (i) carry out a one-time projector calibration, which in our approach calls for additionally calibrating a camera facing downward to the scene plane and includes recovery of the scene plane as a convenient side effect. Next, we use the relative camera-projector-scene plane geometry to (ii) compute a plane-induced homography that distorts the image to be projected in a manner that it appear undistorted to the viewer, and placed in alignment with the axes of a proxy image of the scene. These two points are treated in Sections~\ref{sec:approach:geometry} and \ref{sec:approach:homography}, respectively.

@@

\begin{figure}
    \centerline{\includegraphics[scale=.35]{images/2d3d.png}}
    \caption{Scene plane (gray) recovered via spatial resection with respect to 2D-3D correspondences obtained using a chessboard pattern (blue); 3D circles pattern points obtained by intersection with the scene plane of back-projections (likewise gray) of circles pattern center points detected in the image plane (red). Note that the points in the figure are the points recovered for the projector calibration image in Figure~\ref{fig:2d}(b), acquired by the downwards-facing camera (frustum of the camera in black, with up vector in red).}
    \label{fig:3d}
\end{figure}

@@

\paragraph{Camera calibration} The recovery of 2D-3D correspondences in support of camera calibration is carried out by relying on a planar calibration surface to automatically identify correspondences between the 3D points on the calibration surface and their 2D correspondences in the image plane. The classical calibration surface is a chessboard pattern. The 3D corner points of the chessboard are obtained \textit{a priori} in a coordinate system defined in the plane of the chessboard\footnote{E.g., $(0,0,0), (1.5,0,0), (3,0,0), \dots, (9,7.5,0)$ for a chessboard with $7~\times{}~6$ corners, with each square of length and width of 1.5 unit, respectively. Note that the units of the chessboard's 3D points give the units of the camera calibration, and---owing to how our projector calibration relies on the camera calibration---of the projector calibration as well.}, requiring knowledge only of the dimensions of the chessboard pattern and of the length of a side of a chessboard square. The corresponding 2D points are obtained, in the same order, using a specialized algorithm \cite{bradski2000opencv}. A set of calibration images is acquired, each with the calibration pattern visible in a different part of the image plane, and such that the center and all corners and edges of the image plane are covered, the camera's autofocus setting is off, and the camera's zoom factor remain fixed. 2D-3D correspondences are then recovered for each calibration image, and the resulting list is passed on as input to an optimization procedure that relies on bundle adjustment to yield the camera calibration matrix~$\mathtt{K}_\text{cam}$ and the associated lens distortion model coefficients.

@@

If the projector is calibrated and its pose relative to the scene plane is known, a `virtual' projector (with the same calibration $\mathtt{K}$ and lens distortion coefficients) can be placed elsewhere relative to the scene plane. If we for a moment imagine that the projector---at its recovered pose---functions as a camera,\footnote{Recall that the calibration matrix~$\mathtt{K}$ enables computing both (i) the projection of a scene point to the image plane (the function of a camera), or (ii) the back-projection of a pixel in the image plane, giving a ray into the scene (along which a projector illuminates the scene with the given pixel).} then (i) projecting an image to the scene plane \textit{from the viewpoint of the virtual projector} and (ii) acquiring the resulting projected image from the viewpoint of the recovered projector gives the desired corrective warp. Projecting an image warped in this manner to the scene plane \textit{from the viewpoint of the recovered projector} then has the same effect as projecting the unwarped image to the scene plane from the viewpoint of the virtual projector. This warp can computed using a plane-induced homography, computed analytically as a function of the scene plane, the projector, and the virtual projector.

\begin{figure}
    \centering
    \subfloat[\centering Oblique view.]{{\includegraphics[height=4.2cm]{images/3.png}}}
    \qquad
    \subfloat[\centering Downwards view.]{{\includegraphics[height=4.2cm]{images/3_above3.png}}}
    \caption{The virtual camera is obtained by rotating the camera (top left, black) about its center of projection such that its optical axis be made parallel with the normal vector of the scene plane. The virtual projector is obtained by (i) rotating the projector (bottom left, black) about the point of intersection of its optical axis with the scene plane such that the optical axis be made parallel with the scene plane's normal vector, (ii) rotating the $X$- and $Y$-axes to align them with those of the virtual camera, and (iii) optionally translating along the normal direction to achieve the desired projected image dimensions.} %The virtual projector is thus rendered fronto-parallel with the ground plane and axially aligned with the virtual camera.
    \label{fig:virtualproj}
\end{figure}

\paragraph{Virtual projector} The placement of the virtual projector determines from which pose the image to be projected is to \textit{appear} to have been projected. This placement is produced according to a small handful of steps. First, we (i) rotate the camera about its center of projection to align its optical axis with the normal vector of the scene plane, giving a virtual camera likewise facing directly\footnote{A physical camera placed to face downwards is likely to not face downwards precisely; in contrast, the virtual camera's optical axis is aligned exactly with the scene plane's normal vector, rendering it fronto-parallel with respect to the scene plane.} downwards to the scene plane (cf.\ Figure~\ref{fig:virtualproj}). Next, we (ii) intersect the scene plane with the optical axis (i.e., the ray from the projector's center of projection through the center of the image plane); then, we (iii) rotate the projector's placement about that point of intersection, aligning the optical axis with the scene plane's normal vector and giving an initial virtual projector. Finally, we (iv) align the $X$- and $Y$-axes of the initial virtual projector with those of the virtual camera, which gives the virtual projector (cf.\ again Figure~\ref{fig:virtualproj}). The virtual projector is thus rendered fronto-parallel with the scene plane, enabling projection to the scene plane absent of projective distortions. Optionally, we additionally adjust the height above the scene plane of the virtual projector, in order to satisfy desired projected image dimensions provided in metric units.

@@

We evaluate our approach by augmenting 15 locations across the floorspace at the Pilotfabrik\footnote{\url{https://www.pilotfabrik.at/}} of TU Vienna, a collaborative space for research on Industry 4.0 topics situated in Vienna, Austria. We contrast our approach with two variants of a baseline approach involving manual keystone correction (`Baseline A' and `Baseline B'), by aiming to place an image for each location aligned with the principal axes of the floorspace, absent projective distortions, and with the same dimensions (50~cm~$\times$~31.25~cm). Using our approach to carry out keystone correction analytically, we carried out all steps within a time budget of 15 min. Accordingly, we set 15 min to be the time budget of the first variant of the baseline approach (`Baseline A'), in which we carry out a \texit{manual} keystone correction using the software bundled with the steerable mirror system mounted alongside our projector. The second variant of the baseline approach (`Baseline B') is to again use the bundled software to carry out keystone correction manually, but to record how much time is necessary to produce augmentations within a mean tolerance of XX. All experiments were carried out jointly by the same experienced technician and assistant.

The hardware setup employed in the evaluation comprised a Panasonic PT-RZ660BE projector with a steerable mirror system---used to point the projection to each of the 15 locations---manufactured by Dynamic Projection Institute \cite{rupprecht2020information}. The steerable mirror system was bundled with the MDC-X software for steering the mirror, loading imagery, and optionally carrying out manual keystone correction, such that each position and (warped) image can be registered as a preset. In addition, we used a downward-facing Zed 2 stereo camera manufactured by Stereolabs, yet relied only on the left view. The floorspace used for our experiments measured dimensions of ca.\ 9~m~$\times$~4~m; the projector was mounted at approximately the center of this space, at a height of ca.\ 3.5 m.

\paragraph{Our approach} We began by carrying out a one-time\footnote{Note that once the camera is calibrated, that calibration can be reused if the camera's intrinsics remain fixed, i.e., no change is made to the zoom factor.} calibration of the camera, acquiring 10 camera calibration images (cf.\ Section~\ref{sec:approach:geometry}) and feeding the images as input to our camera calibration module. Separately, for each of the 15 target locations, we produced a projector calibration image (cf.\ again Section~\ref{sec:approach:geometry}) by projecting the circles pattern image to the location in question using the steerable mirror, placing a checkerboard pattern beside the projected pattern, and acquire the image using the downward-facing camera. We then fed these images alongside the output of the camera calibration module to our projector calibration module. For each of the target locations, the steerable mirror was made to point to that location, the circles pattern image was projected to the scene plane, an image was using the camera, and the location was registered in the MDC-X software as a preset. The output of the projector calibration module is a homography per input projector calibration image (cf.\ Section~\ref{sec:approach:homography}). Next, we warped the images to be projected to the respective locations using their corresponding homography, using a third dedicated custom module. These warped images were finally imported into the MDC-X software and associated with their respective location presets.

The total amount of time to carry out all the above steps amounted to ca.\ 15 min, with ca.\ 2 min going to acquisition of the camera calibration images, and ca.\ 5 min going to that of projector calibration images. The remainder of the time was spent running our modules or working with the MDC-X software.
